Extreme Digitisation For Ground-Based Cosmic Microwave Background Experiments

### 2018-12-21 digitisation.tex
random

This work is structured as follows. In section \S\ref{sec:dig} we detail the challenges that come with handling large data volumes, introduce the process of extreme digitisation and lay out the framework used to test its performance. We present results for white and $1/f$ detector noise in section \S\ref{sec:results}. We summarise our findings in \S\ref{sec:conclusions}.

@@

The deduced additional noise for one, two, and three bit digitisation is shown in Table \ref{tab:extranoisewhite} in the appendix. We see that 3 bit digitisation performs best, followed by two bit and one bit. This is expected, since more bits allow a more faithful representation of the input signal. We observe that the fractional increase, $\Delta \sigma / \sigma$, is independent of the hits per pixel in the maps and that the compression performs as well for polarisation as is does for temperature data. \changed{Finally, it is striking how little noise is added. On average one-bit digitisation leads to a $(25.13\pm 0.94)\%$ increase in the map noise level, two-bit yields a $(6.40\pm0.41)\%$ increase and three-bit adds only $(1.76\pm0.22)\%$ ($1\sigma$ confidence intervals given). This is impressive, given that all schemes considered reduce the TOD volume by at least an order of magnitude.}

@@

\begin{figure}[htb]\centering
\includegraphics[width=0.5\textwidth,clip]{diffwn.pdf}
  \caption[Current ]{
  \changed{(FIGURE HAS BEEN CHANGED) TT, EE, and BB power spectra (upper, middle and lowest panel respectively) of difference maps for one-bit (green marker, high), two-bit (orange marker, medium), and three-bit digitisation (violet marker, low). The control case using 64-bit TOD is also shown (black marker). All power spectra are rebined to $\Delta \ell = 200$. The vertical axes are logarithmic. The noise increase due to digitisation appears to be independent of angular scale. This plot shows results for a simulation with $1,024,000$ hits per map pixel using Gaussian white detector noise.}
\label{fig:diffpswn}
}
\end{figure}

@@

While digitisation clearly works extremely well for white noise, real ground-based CMB experiments face several sources of low-frequency noise. \changed{High on the list is atmospheric noise. Modelling the atmosphere as a thin sheet dominated by Kolmogorov turbulence, we can describe fluctuations with the noise profile \citep{lay2000}:}

@@

We recover similar results to the white noise case, with three bit digitisation performing the best, followed by two bit, and one bit. Table \ref{tab:extranoiseoof} in the appendix displays the fractional increase to the detector noise level due to digitisation, $\Delta \sigma / \sigma$. While polarisation observations seem to incur a slightly higher noise penalty compared to temperature, deviations do not exceed $3\sigma$. \changed{On average one-bit digitisation leads to a $(24.4\pm 1.4)\%$ increase in the map noise level, two-bit adds $(6.21\pm0.50)\%$, and three-bit yields an increase of only $(1.71\pm0.21)\%$ ($1\sigma$ confidence intervals given). Thus the effectiveness of few bit digitisation is not limited to the white noise case, but \changed{extends} to more realistic noise profiles.}

@@

\subsection{\changed{Existing Compression Techniques}}

\changed{Current-generation ground-based CMB experiments employ lossless and lossy compression techniques to manage their transmission bottlenecks. To test whether few-bit digitisation is competitive to existing lossy compression techniques, it is necessary to see if it still provides a substantial reduction in data volume if used in conjunction with lossless compression algorithms.}

>>>>

This work is structured as follows. In section \S\ref{sec:dig} we detail the challenges that come with handling large data volumes, introduce the process of extreme digitisation and lay out the framework used to test its performance. 
We present results for white and $1/f$ detector noise in section \S\ref{sec:results}. 
We summarise our findings in \S\ref{sec:conclusions}.
\changed{We report uncertainties based on the symmetric 68\% confidence regions. }

@@

The deduced additional noise for one, two, and three bit digitisation is shown in Table \ref{tab:extranoisewhite} in the appendix. 
We see that 3 bit digitisation performs best, followed by two bit and one bit. 
This is expected, since more bits allow a more faithful representation of the input signal. 
We observe that the fractional increase, $\Delta \sigma / \sigma$, is independent of the hits per pixel in the maps and that the compression performs as well for polarisation as is does for temperature data. \changed{Finally, it is striking how little noise is added. 
On average one-bit digitisation leads to a $(25.13\pm 0.94)\%$ increase in the map noise level, two-bit yields a $(6.40\pm0.41)\%$ increase and three-bit adds only $(1.76\pm0.22)\%$. 
This is impressive, given that all schemes considered reduce the TOD volume by at least an order of magnitude.}

@@

\begin{figure}[htb]\centering
\includegraphics[width=0.5\textwidth,clip]{diffwn.pdf}
  \caption[Current ]{
  \changed{TT, EE, and BB power spectra (upper, middle and lowest panel respectively) of difference maps for one-bit (green marker, high), two-bit (orange marker, medium), and three-bit digitisation (violet marker, low). The control case using 64-bit TOD is also shown (black marker). All power spectra are rebined to $\Delta \ell = 200$. The vertical axes are logarithmic. The noise increase due to digitisation appears to be independent of angular scale. This plot shows results for a simulation with $1,024,000$ hits per map pixel using Gaussian white detector noise.}
\label{fig:diffpswn}
}
\end{figure}

@@

\changed{While digitisation clearly works extremely well for white noise, real experiments face several sources of low-frequency noise. 
High on the list for ground-based experiments is atmospheric noise. 
For typical observing conditions in Antarctica, the atmosphere turbulence is well-described by Kolmogorov turbulence in the thin-sheet regime \citep{bussmann2005}. 
In this regime, we can describe fluctuations with the noise profile \citep{lay2000}:}

@@

We recover similar results to the white noise case, with three bit digitisation performing the best, followed by two bit, and one bit. Table \ref{tab:extranoiseoof} in the appendix displays the fractional increase to the detector noise level due to digitisation, $\Delta \sigma / \sigma$. While polarisation observations seem to incur a slightly higher noise penalty compared to temperature, deviations do not exceed $3\sigma$. \changed{On average one-bit digitisation leads to a $(24.4\pm 1.4)\%$ increase in the map noise level, two-bit adds $(6.21\pm0.50)\%$, and three-bit yields an increase of only $(1.71\pm0.21)\%$. 
The fractional added noise from few-bit digitisation is consistent with being identical for the white noise and realistic noise profiles.}

@@

\subsection{\changed{Interaction with Existing Compression Techniques}}

\changed{Current ground-based CMB experiments already employ 
 lossless and lossy compression techniques to manage their transmission bottlenecks. To test whether few-bit digitisation is competitive to existing lossy compression techniques, it is necessary to see if it still provides a substantial reduction in data volume if used in conjunction with lossless compression algorithms.}
